Constraint Algebras \\ of Multisymplectic Observables

### 2025-01-20 msconstraints-short-main.tex
Conference version.

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%	
\subsection{Regular reduction} 
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%	
\begin{frame}{Regular reduction in symplectic geometry}
	\textbf{\color{UniGreen}Symplectic reduction:}~~
	\\ 
	{\it \small
	Procedure associating to any (suitably regular) pair of symplectic manifold and Hamiltonian action another symplectic manifold of smaller dimension.}
	\vfill

	\pause
	\begin{columns}
		\begin{column}{0.40\textwidth}
			\includegraphics[width=\textwidth]{Pictures/Reduction}
		\end{column}

		\begin{column}{0.6\textwidth}
				\textbf{\color{UniGreen}In mechanics:}~~
				\\
			{\it \small
				it embodies the process of restricting the dynamics of the system to the level sets of the conserved quantities pertaining to the symmetry group.		
			}
			\\[.2em]
			\color{gray}\small( e.g. restricting to studying a point-like particle in a central potential by studying it in radial coordinates)
		\end{column}	
	\end{columns}	
\end{frame}
\note[itemize]{
	\item  In classical mechanics symmetry reduction plays an important role. Mathematically, this is usually phrased in terms of Marsden-Weinstein reduction
	\item {Symplectic Reduction} takes two inputs:
	\\ 1. Hamiltonian $G$-space $(M,\omega,{\color{orange}G},{\color{blue}\mu})$
	\\ 2. parameter ${\color{blue}\phi}\in\g^*$
	\item The \emph{reduced space} is $M_\phi:={\color{blue}\mu^{-1}(\phi)}{\color{orange}/G_\phi}$.
	\item (Marsden--Weinstein '74, Meyer '73) THM:\\
		If $\mu^{-1}(\phi)\subset M$ is smooth and $G_\phi\curvearrowright \mu^{-1}(\phi)$ is free and proper, then there is a \textbf{unique symplectic form} $\omega_\phi\in\Omega^2(M_\phi)$ such that $\pi^*\omega_\phi = i^*\omega$.

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Singular reduction}
\begin{frame}{Symplectic singular reduction schemes}
	\begin{block}{The gist of singular reduction}
		 \begin{itemize}
			 \item[-] when $\mu$ is singular (i.e. $\mu^{-1}(0)$ is not a mfd.), the (geometrically) reduced space may not exist.
			 \item[-] a \emph{singular reduction scheme} is a procedure to construct a "reduced" algebra of observable out of the given data
			 \item[-] such that it corresponds to the algebra of observable of the reduced manifold in the regular case.
		 \end{itemize}
	\end{block}
	 %
	 \pause
	 %
	\begin{thmblock}[Sniatycki-Weinstein reduction \cite{SniatyckiWeinstein83}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
			Given: & $(M,\omega)$ symplectic
			\\
			& $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
			\\[.4em]
			Then: & 
			$\displaystyle \left[\sfrac{C^\infty(M)}{I_\mu}\right]^G$
			admits a Poisson algebra structure 
			\footnote{$I_\mu$ = associative ideal generated by $\widetilde{\mu}(\g)$}			
			\\
			& it agrees with the M--W reduction in the regular case.		
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}
 \end{frame}
\note[itemize]{
 \item Let $(M,\omega,G,\mu)$ be a symplectic Hamiltonian $G$-space.
 \item The \emph{momentum ideal} is the associative ideal $I_\mu\subset C^\infty(M)$ generated by the momenta $\mu_{\xi}$ for any $\xi\in \mathfrak{g}$.
		Namely
		\[
			I_\mu 
			= 
			\Big\langle \mu_\xi \Big\rangle_{\xi \in \mathfrak{g}}^{\text{\tiny asso.}} 
			=
			\left\{
				\left.
					\sum_{i=1}^n f_i ~ \mu_{\xi_i}
				\right|~
					f_i \in C^\infty(M),~ \xi_i\in \mathfrak{g},~  1 \leq i \leq n
			\right\}
		\]

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Constraints Algebras}
\begin{frame}[fragile]{Attempting algebraic reduction}
When the action is not 'free and proper', we can try to mimic this directly using the function algebras:
% https://q.uiver.app/#q=WzAsNixbMCwwLCJNIl0sWzEsMCwiXFxtdV57LTF9KDApPVMiXSxbMywwLCJNX3IiXSxbMCwxLCJDXlxcaW5mdHkoTSkiXSxbMSwxLCJDXntcXGluZnR5fShTKT1cXGZyYWN7Q157XFxpbmZ0eX0oTSl9e0lfe1N9fSJdLFszLDEsIkNee1xcaW5mdHl9KE1fcik9Q15cXGluZnR5KFMpXkciXSxbMSwyLCJcXHN1YnN0YWNre3F1b3RpZW50fSIsMCx7InN0eWxlIjp7ImhlYWQiOnsibmFtZSI6ImVwaSJ9fX1dLFszLDQsIlxcc3Vic3RhY2t7cXVvdGllbnR9IiwwLHsic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoiZXBpIn19fV0sWzAsMSwiXFxzdWJzdGFja3tyZXN0cmljdGlvbn0iLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFs0LDUsIlxcc3Vic3RhY2t7cmVzdHJpY3Rpb259IiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XV0=
\[\begin{tikzcd}
	M & {\mu^{-1}(0)=S} && {M_r} \\
	{C^\infty(M)} & {C^{\infty}(S):=\frac{C^{\infty}(M)}{I_{S}}} && {C^{\infty}(M_r)=C^\infty(S)^G}
	\arrow["{\substack{restriction}}", squiggly, from=1-1, to=1-2]
	\arrow["{\substack{quotient}}", two heads, from=1-2, to=1-4]
	\arrow["{\substack{quotient}}", two heads, from=2-1, to=2-2]
	\arrow["{\substack{restriction}}", squiggly, from=2-2, to=2-4]
\end{tikzcd}\]

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Constraint Algebras} 
	\begin{block}{Idea:}
	    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
			\item[-] Both regular/geometric and singular/algebraic reduction are 2-steps procedures (1. restrict 2. quotient).
			\item[-] Let's work in a category where all data needed for the reduction are bundled together in suitable categorical objects. {\tiny (To avoid pathologies arising from these steps)}
	    \end{itemize}
	\end{block}
	\pause
	\vfill
	\begin{defblock}[Constraint algebra triples \cite{Dippell2023}]
		A (strong embedded) constraint algebra $\mathcal A$ is given by a triple $(\mathcal A_T, \mathcal A_N, \mathcal A_0)$, where
		\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
			\item[•] $\mathcal A_T$ is an algebra \emph{("total space")}
			\item[•] $\mathcal A_N\subset \mathcal A_T$ a subalgebra  \emph{("normal space")}
			\item[•] $\mathcal A_0\subset \mathcal A_N$ a two-sided ideal in $\mathcal A_T$ \emph{("zero space")}
		\end{itemize}
	\end{defblock}
	\vfill
	\pause
	\begin{exblock}
		Let $S$ be a closed set. Algebraically, we describe $C^\infty(S)$ by $\frac{C^\infty(M)}{I_S}$.
		\\ 
		We can think of it as the constraint triple:
		$$
			\mathcal A=(\mathcal A_T,\mathcal A_N,\mathcal A_0)=(C^\infty(M),C^\infty(M), I_S)
		$$
	\end{exblock}

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Other approaches to reduction}

\begin{table}[]
	\begin{tabular}{l|l|l||l}
		algebra $\mathcal{A}_T$ & subalgebra $\mathcal{A}_N$ & ideal $\mathcal{A}_0$ & comment \\
		\hline

@@

	$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_S)^G$   &  $\substack{C^\infty(M)^G=\{f~|~gf=f\forall g\in G\}}$ \\
	\hline \pause
		$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_\mu)^G$   & $\substack{I_\mu=\{\sum_{i=1}^k f_i\langle \mu,\xi_i\rangle|~f_i\in C^\infty(M), \xi_i\in \mathfrak g\}}$\\
		\hline \pause
	$C^\infty(M)$&      $N_\mu$     &   $I_\mu$   & $\substack{N_\mu= \{ f | gf-f\in I_\mu \forall g\in G\}}$\\
	\hline

	\end{tabular}\pause
\end{table}
	\vfill
{\bf Remarks }names in order: 'naive' , Dirac, Sniatycki, Arms-Cushman-Gotay, Sniatycki-Weinstein.\pause \\
When $G$ connected $N_\mu=N(I_\mu):=\{f~|~\{f,I_\mu\}\subset I_\mu\}$. 
\end{frame}
\note[itemize]{
	\item  
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

@@

	\vfill
	\begin{table}[]
		\begin{tabular}{l|l|l}
		 geometry & algebra & translation \\
			\hline	
		smooth manifold $M$ & comm. algebra $\mathcal A_T$  	& $\mathcal A_T=C^{\infty}(M)$\\ \hline\pause
		vector fields $\mathfrak X(M)$ & derivations $\Der(\mathcal A_T)$ &  $\Der(C^\infty(M))=\mathfrak X(M)$\\ \hline\pause
		closed subset $S\subset M$ & ideals $\mathcal A_0\subset \mathcal A_T$ & $\mathcal A_0=I_S=\{f~|~f|_S=0\}$ \\ \hline\pause
		action $G\circlearrowleft M$ & Lie algebra map & infinitesimal generators\\
		 of connected group &$\underline{\cdot}:\mathfrak g\to \Der(\mathcal A_T)$ & of the action \\ \hline \pause
		 $G$ action preserves $S$ & $\underline \xi(I_S)\subset I_S$ $\forall \xi\in\mathfrak g$& \color{red} Careful! Not equivalent! 
		 \\ \hline
		\end{tabular}\pause
	\end{table}
	\vfill

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Higher Observables Constraint Algebra}
\begin{frame}[fragile]{Higher Observables Constraint Algebra}
	\begin{defpropblock}[$L_\infty$-observables constraint algebra]
		The following $L_\infty$ algebras form a constraint triple
		\begin{align*}
			\Ham_T =&~ L_\infty(M,\omega) \\
			\Ham_N =&~
			{\small
			\left\lbrace
			\begin{array}{l}
				\alpha\in \Omega^{\leq n-2}(M) ~\mathrm{, respectively}\\	
				(\alpha,X) \in \Omega^{n{-}1}(M) \oplus\mathfrak X(M)
			\end{array}
			~\left\vert~
			\begin{array}{l l}
				\iota_\xi \alpha \in (\Lambda \mathfrak L)^*_0 &\\
				L_\xi\alpha \in (\Lambda \mathfrak L)^*_0&~\forall \xi \in F\\
				\hline
				\iota_X \omega = -d \alpha &\\
				X(I_S)\subset I_S &\\
				L_\xi X \in C^{\infty}(M) F + I_S\mathfrak X(M)
				&~\forall \xi \in F\\
			\end{array}
			\right.
			\right\rbrace 
			}
			\\
			\Ham_0 =&~
		{\small
			\left\lbrace
			\begin{array}{l}
				\alpha\in \Omega^{\leq n-2}(M) ~\mathrm{, respectively}\\	
				(\alpha,X) \in \Omega^{n{-}1}(M) \oplus \mathfrak X(M)
			\end{array}
			~\left\vert~
			\begin{array}{l}
				\alpha \in (\Lambda \mathfrak L)^*_0\\
				\hline
				\iota_X \omega = -d \alpha\\
				X \in C^{\infty}(M) F + I_S\mathfrak X(M)
			\end{array}
			\right.
			\right\rbrace
		}
		\end{align*}
		where $(\Lambda^k \mathfrak L)^*_0=\{\alpha~|~\alpha(v_1,...,v_k)\in \mathcal A_0~\forall v_i\in \mathfrak L_N\}$.
	\end{defpropblock}
	\vfill
	\pause
	Hence, The following subquotient $\frac{\Ham_N}{\Ham_0}$ of $\Ham_T$ forms an $L_\infty$-algebra.
\end{frame}
\note[itemize]{
	\item 	\begin{defblock}[Reducible v.fields ]
			\begin{displaymath}
				\X(M)_{[N]} :=
				\left\lbrace
					v \in \X(M)
				~\left\vert~
					\begin{array}{l}
						\L_v (I_N) \subseteq I_N	\\		
						\L_v (\X_g) \subseteq \X_g + I_\X
					\end{array}
				\right\rbrace\right.
			\end{displaymath}
%			\blfootnote{
%			 $\X_g$ = $C^\infty(M)$-module generated by the fundamental distribution.
%			}

	\end{defblock}	
	%
	\pause
	%
	\begin{defblock}[Reducible forms ]
		\begin{displaymath}
			\Omega(M)_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)
			~\left\vert~
				\begin{array}{l r}
					\L_\underline{\xi} \,\alpha \in I_{\Omega(N)}	\\		
					\iota_\underline{\xi} \,\alpha \in I_{\Omega(N)}	& \forall \xi \in \g				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}	
	%
	\pause
	%
	\begin{defblock}[Reducible Hamiltonian forms]
		\begin{displaymath}
			(\Omega(M)_{ham}^{n-1})_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)_{ham}^{n-1}
			~\left\vert~
				\begin{array}{l r}
					\alpha \text{ is a reducible form} \\
					\vHam_\alpha \text{ is a reducible v.field}
				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}	
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Conclusions}
\begin{frame}{Lessons learned}
	\begin{itemize} 
\item Constraint triples permit to do differential geometry/ Cartan calculus on singular subquotients,\\ ... in particular constructing observable algebras on them.\pause
\item Even if $\omega$ is non-degenerate, $\omega_r$ could degenerate, forcing us to pull along the Hamiltonian vector fields explicitly\pause
\item Coomentum maps in the sense of the Leibniz structure seem better for reducing than $L_\infty$-comomenta \pause
\item The algebraic approach can treat algebroid / singular foliation symmetries and remembers vanishing orders a the singularities. \pause
\item The reduction seems to give something novel even in the symplectic case \pause
\item A 'softer' notion of constraint $L_\infty$ algebra is necessary to obtain all observables of the reduced space.
	\pause
			\vfill
		  \centering 
		  {\Huge\color{red} 
		  \emph{Thank you for your attention!}}
			\vfill
	\end{itemize}
\end{frame}
\note[itemize]{
	\item 
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

@@

\subsection{Preview of our answer}
\begin{frame}
We will develop a multisymplectic reduction in this talk, symplectically it is:
\begin{block}{Construction: Multisymplectic reduction in the symplectic case}
Let $Q_S=\{f~|~ f|_S=0\rm{~and~} df|_S=0\}$ Then the $L_\infty$-reduction is:
$$
\frac{N(I_S)\cap N(I_\mu+Q)}{I_\mu+Q}
$$
\end{block}	\pause
Such a formula can be more generally applied to a Poisson algebra $P$ with associative ideals $I\subset J$ such that $\{I,I\}\subset I$ and $\{I,J\}\subset J$:
$$
\frac{N(J)\cap N(I+Q)}{I+Q}
$$
with $Q=\{f\in J|~\{f,P\}\subset J\}$
\end{frame}

>>>>

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%	
\subsection{Regular reduction} 
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%	
\begin{frame}{Regular reduction in symplectic geometry}

	\textbf{\color{UniGreen}Symplectic reduction:}~~
	\\ 
	{\it \small
	Procedure associating to any (suitably regular) pair of symplectic manifold and Hamiltonian action another symplectic manifold of smaller dimension.}
	\vfill

	\pause

	\begin{center}
		\includegraphics[width=.5\textwidth]{Pictures/Reduction}	
	\end{center}

	\textbf{\color{UniGreen}In mechanics:}~~
				\\
			{\it \small
				it embodies the process of restricting the dynamics of the system to the level sets of the conserved quantities pertaining to the symmetry group.		
			}
			\\[.2em]
			\color{gray}\small( e.g. restricting to studying a point-like particle in a central potential by studying it in radial coordinates)

\end{frame}
\note[itemize]{
	\item  In classical mechanics symmetry reduction plays an important role. Mathematically, this is usually phrased in terms of Marsden-Weinstein reduction
	\item {Symplectic Reduction} takes two inputs:
	\\ 1. Hamiltonian $G$-space $(M,\omega,{\color{orange}G},{\color{blue}\mu})$
	\\ 2. parameter ${\color{blue}\phi}\in\g^*$
	\item The \emph{reduced space} is $M_\phi:={\color{blue}\mu^{-1}(\phi)}{\color{orange}/G_\phi}$.
	\item (Marsden--Weinstein '74, Meyer '73) THM:\\
		If $\mu^{-1}(\phi)\subset M$ is smooth and $G_\phi\curvearrowright \mu^{-1}(\phi)$ is free and proper, then there is a \textbf{unique symplectic form} $\omega_\phi\in\Omega^2(M_\phi)$ such that $\pi^*\omega_\phi = i^*\omega$.

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Singular reduction}
\begin{frame}{Symplectic singular reduction schemes}
	\begin{block}{The gist of singular reduction}
		 \begin{itemize}
			 \item[-] when $\mu$ is singular (i.e. $\mu^{-1}(0)$ is not a mfd.), the (geometrically) reduced space may not exist.
			 \item[-] a \emph{singular reduction scheme} is a procedure to construct a "reduced" algebra of observable out of the given data
			 \item[-] such that it corresponds to the algebra of observable of the reduced manifold in the regular case.
		 \end{itemize}
	\end{block}
	 %
	 \pause
	 \vfill
	 %
	\begin{thmblock}[Sniatycki-Weinstein reduction \cite{SniatyckiWeinstein83}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
			Given: & $(M,\omega)$ symplectic
			\\
			& $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
			\\[.4em]
			Then: & 
			$\displaystyle \left[\sfrac{C^\infty(M)}{I_\mu}\right]^G$
			admits a Poisson algebra structure 
			\footnote{$I_\mu$ = associative ideal generated by $\widetilde{\mu}(\g)$}			
			\\
			& it agrees with the M--W reduction in the regular case.		
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}
 \end{frame}
\note[itemize]{
 \item Let $(M,\omega,G,\mu)$ be a symplectic Hamiltonian $G$-space.
 \item The \emph{momentum ideal} is the associative ideal $I_\mu\subset C^\infty(M)$ generated by the momenta $\mu_{\xi}$ for any $\xi\in \mathfrak{g}$.
		Namely
		\[
			I_\mu 
			= 
			\Big\langle \mu_\xi \Big\rangle_{\xi \in \mathfrak{g}}^{\text{\tiny asso.}} 
			=
			\left\{
				\left.
					\sum_{i=1}^n f_i ~ \mu_{\xi_i}
				\right|~
					f_i \in C^\infty(M),~ \xi_i\in \mathfrak{g},~  1 \leq i \leq n
			\right\}
		\]

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Constraints Algebras}
\begin{frame}[fragile]{Attempting algebraic reduction naively}
When the action is not 'free and proper', we can try to mimic this directly using the function algebras:
% https://q.uiver.app/#q=WzAsNixbMCwwLCJNIl0sWzEsMCwiXFxtdV57LTF9KDApPVMiXSxbMywwLCJNX3IiXSxbMCwxLCJDXlxcaW5mdHkoTSkiXSxbMSwxLCJDXntcXGluZnR5fShTKT1cXGZyYWN7Q157XFxpbmZ0eX0oTSl9e0lfe1N9fSJdLFszLDEsIkNee1xcaW5mdHl9KE1fcik9Q15cXGluZnR5KFMpXkciXSxbMSwyLCJcXHN1YnN0YWNre3F1b3RpZW50fSIsMCx7InN0eWxlIjp7ImhlYWQiOnsibmFtZSI6ImVwaSJ9fX1dLFszLDQsIlxcc3Vic3RhY2t7cXVvdGllbnR9IiwwLHsic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoiZXBpIn19fV0sWzAsMSwiXFxzdWJzdGFja3tyZXN0cmljdGlvbn0iLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFs0LDUsIlxcc3Vic3RhY2t7cmVzdHJpY3Rpb259IiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XV0=
\[\begin{tikzcd}
	M & {\mu^{-1}(0)=S} && {M_r} \\
	{C^\infty(M)} & {C^{\infty}(S):=\frac{C^{\infty}(M)}{I_{S}}} && {C^{\infty}(M_r)=C^\infty(S)^G}
	\arrow["{\substack{restriction}}", squiggly, from=1-1, to=1-2]
	\arrow["{\substack{quotient}}", two heads, from=1-2, to=1-4]
	\arrow["{\substack{quotient}}", two heads, from=2-1, to=2-2]
	\arrow["{\substack{restriction}}", squiggly, from=2-2, to=2-4]
\end{tikzcd}\]

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Constraint Algebras} 
	\begin{block}{Idea:}
	    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
			\item[-] Both regular/geometric and singular/algebraic reduction are 2-steps procedures (1. restrict 2. quotient).
			\item[-] Let's work in a category where all data needed for the reduction are bundled together in a suitable categorical objects. {\tiny (To avoid pathologies arising from these steps)}
	    \end{itemize}
	\end{block}
	\pause
	\vfill
	\begin{defblock}[Constraint algebra triples \cite{Dippell2023}]
		A (strong embedded) constraint algebra $\mathcal A$ is given by a triple $(\mathcal A_T, \mathcal A_N, \mathcal A_0)$, where
		\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
			\item[•] $\mathcal A_T$ is an algebra \emph{("total space")}
			\item[•] $\mathcal A_N\subset \mathcal A_T$ a subalgebra  \emph{("normal space")}
			\item[•] $\mathcal A_0\subset \mathcal A_N$ a two-sided ideal in $\mathcal A_T$ \emph{("zero space")}
		\end{itemize}
	\end{defblock}
	\vfill
	\pause
	\begin{exblock}
		Let $S$ be a closed set. Algebraically, we describe $C^\infty(S)$ by $\frac{C^\infty(M)}{I_S}$.
		\\ 
		We can think of it as the constraint triple:
		$$
			\mathcal A=(\mathcal A_T,\mathcal A_N,\mathcal A_0)=(C^\infty(M),C^\infty(M), I_S)
		$$
	\end{exblock}

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Other approaches to reduction}
 	Many singular reduction schemes adhere to this pattern!
 	\vfill
 	\pause
\begin{table}[]
	\begin{tabular}{l|l|l||l}
		algebra $\mathcal{A}_T$ & subalgebra $\mathcal{A}_N$ & ideal $\mathcal{A}_0$ & comment \\
		\hline

@@

	$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_S)^G$   &  $\substack{C^\infty(M)^G=\{f~|~gf=f\forall g\in G\}}$ \\
	\hline \pause
		$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_\mu)^G$   & $\substack{I_\mu=\{\sum_{i=1}^k f_i\langle \mu,\xi_i\rangle|~f_i\in C^\infty(M), \xi_i\in \mathfrak g\}}$\\
		\hline \pause
	$C^\infty(M)$&      $N_\mu$     &   $I_\mu$   & $\substack{N_\mu= \{ f | gf-f\in I_\mu \forall g\in G\}}$ \footnote{When $G$ connected $N_\mu=N(I_\mu):=\{f~|~\{f,I_\mu\}\subset I_\mu\}$.}\\
	\hline

	\end{tabular}\pause
\end{table}
	\vfill
	{\bf Remarks }names in order: 'naive' , Dirac, Sniatycki, Arms-Cushman-Gotay, Sniatycki-Weinstein.

@@

	\vfill
	\begin{table}[]
		\begin{tabular}{l|l|l}
		 geometry & algebra & translation \\
			\hline	
		smooth manifold $M$ & comm. algebra $\mathcal A_T$  	& $\mathcal A_T=C^{\infty}(M)$\\ \hline\pause
		vector fields $\mathfrak X(M)$ & derivations $\Der(\mathcal A_T)$ &  $\Der(C^\infty(M))=\mathfrak X(M)$\\ \hline\pause
		closed subset $S\subset M$ & ideals $\mathcal A_0\subset \mathcal A_T$ & $\mathcal A_0=I_S=\{f~|~f|_S=0\}$ \\ \hline\pause
		action $G\circlearrowleft M$ & Lie algebra map & infinitesimal generators\\
		 of connected group &$\underline{\cdot}:\mathfrak g\to \Der(\mathcal A_T)$ & of the action \\ \hline \pause
		 $G$ action preserves $S$ & $\underline \xi(I_S)\subset I_S$ $\forall \xi\in\mathfrak g$& \color{red} Careful! Not equivalent! 
		 \\ \hline
		\end{tabular}
	\end{table}
	\vfill

@@

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Higher Observables Constraint Algebra}
\begin{frame}[shrink]{Higher Observables Constraint Algebra}
	\begin{defpropblock}[$L_\infty$-observables constraint algebra]
		The following $L_\infty$ algebras form a constraint triple
		\begin{align*}
			\Ham_T =&~ L_\infty(M,\omega) \\[1em]
			\Ham_N =&~
			{\small
			\left\lbrace
			\begin{array}{l}
				\alpha\in \Omega^{\leq n-2}(M) ~\mathrm{, respectively}\\	
				(\alpha,X) \in \Omega^{n{-}1}(M) \oplus\mathfrak X(M)
			\end{array}
			~\left\vert~
			\begin{array}{l l}
				\iota_\xi \alpha \in \Omega_0 &\\
				\L_\xi\alpha \in \Omega_0&~\forall \xi \in F\\
				\hline
				\iota_X \omega = -d \alpha &\\
				X(I_S)\subset I_S &\\
				\L_\xi X \in C^{\infty}(M) F + I_S\mathfrak X(M)
				&~\forall \xi \in F\\
			\end{array}
			\right.
			\right\rbrace 
			}
			\\[1em]
			\Ham_0 =&~
		{\small
			\left\lbrace
			\begin{array}{l}
				\alpha\in \Omega^{\leq n-2}(M) ~\mathrm{, respectively}\\	
				(\alpha,X) \in \Omega^{n{-}1}(M) \oplus \mathfrak X(M)
			\end{array}
			~\left\vert~
			\begin{array}{l}
				\alpha \in \Omega_0\\
				\hline
				\iota_X \omega = -d \alpha\\
				X \in C^{\infty}(M) F + I_S\mathfrak X(M)
			\end{array}
			\right.
			\right\rbrace
		}
		\end{align*}
		where: 
			\begin{align*}
				\small\Omega_0^k &:=~\{\alpha \in \Omega^k(M)~|~\alpha(v_1,...,v_k)\in  I_S~\forall v_i\in \X_N\} 
				\\
				\mathfrak{X}_N &:=~ 
				\left\lbrace
					v \in \X(M)
				~\left\vert~
					\begin{array}{l}
						\L_v (I_N) \subseteq I_N	\\		
						\L_v (C^{\infty}(M) F) \subseteq C^{\infty}(M) F + I_S\mathfrak X(M)
					\end{array}
				\right\rbrace\right.
			\end{align*} 
	\end{defpropblock}
	\vfill
	\pause
	Hence, the subquotient $\frac{\Ham_N}{\Ham_0}$ of $\Ham_T$ forms an $L_\infty$-algebra.
\end{frame}
\note[itemize]{
	\item 	\begin{defblock}[Reducible v.fields ]
			\begin{displaymath}
				\X(M)_{[N]} :=
				\left\lbrace
					v \in \X(M)
				~\left\vert~
					\begin{array}{l}
						\L_v (I_N) \subseteq I_N	\\		
						\L_v (\X_g) \subseteq \X_g + I_\X
					\end{array}
				\right\rbrace\right.
			\end{displaymath}
			$\X_g$ = $C^\infty(M)$-module generated by the fundamental distribution.
	\end{defblock}	
	%
	\pause
	%
	\begin{defblock}[Reducible forms ]
		\begin{displaymath}
			\Omega(M)_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)
			~\left\vert~
				\begin{array}{l r}
					\L_\underline{\xi} \,\alpha \in I_{\Omega(N)}	\\		
					\iota_\underline{\xi} \,\alpha \in I_{\Omega(N)}	& \forall \xi \in \g				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}	
	%
	\pause
	%
	\begin{defblock}[Reducible Hamiltonian forms]
		\begin{displaymath}
			(\Omega(M)_{ham}^{n-1})_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)_{ham}^{n-1}
			~\left\vert~
				\begin{array}{l r}
					\alpha \text{ is a reducible form} \\
					\vHam_\alpha \text{ is a reducible v.field}
				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}	
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize} 
\item Constraint triples permit to do differential geometry/ Cartan calculus on singular subquotients,\\ ... in particular constructing observable algebras on them.\pause
\item Even if $\omega$ is non-degenerate, $\omega_r$ could degenerate, forcing us to pull along the Hamiltonian vector fields explicitly\pause
\item Coomentum maps in the sense of the Leibniz structure seem better for reducing than $L_\infty$-comomenta \pause
%\item The algebraic approach can treat algebroid / singular foliation symmetries and remembers vanishing orders a the singularities. \pause
\item The reduction seems to give something novel even in the symplectic case 
\begin{block}{Construction: Multisymplectic reduction in the symplectic case}
Let $Q=\{f~|~ f|_S=0\rm{~and~} df|_S=0\}$ Then the $L_\infty$-reduction is:
$$
\frac{N(I_S)\cap N(I_\mu+Q)}{I_\mu+Q}
$$
\end{block}

	\pause
			\vfill
		  \centering 
		  {\Huge\color{red} 
		  \emph{Thank you for your attention!}}
			\vfill
	\end{itemize}
\end{frame}
\note[itemize]{
	\item We will develop a multisymplectic reduction in this talk, symplectically is (see slide).
	\item Such a formula can be more generally applied to a Poisson algebra $P$ with associative ideals $I\subset J$ such that $\{I,I\}\subset I$ and $\{I,J\}\subset J$:
$$
\frac{N(J)\cap N(I+Q)}{I+Q}
$$
with $Q=\{f\in J|~\{f,P\}\subset J\}$

@@
